### 2024-06-07 confoundingSimulations.qmd
new ms repo, revised ms from april 2024 data-cleaning version

---
title: Confounding Simulations
output: html_document
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
page-layout: full
format:
  html:
    code-fold: false
    code-summary: "Show the code"
---

```{r}
#| label: setup

library(tidyverse)
library(broom)
```

First let's set up the true population effects for X (gaming playtime) on Y (wellbeing), which are in turn both exogenously caused by C (confounds).

```{r}
#| label: specify-true-effects

n <- 500
sims <- 1000
eff_xy <- .2
eff_cx <- .5
eff_cy <- -.5
```

Now let's run 1000 studies with these effects and see.

```{r}
#| label: simulate-studies

simulate_study <- function(n = 500,
                           sims = 1000,
                           eff_xy = .2,
                           eff_cx = .5,
                           eff_cy = -.5, 
                           controlForConfound = FALSE) {
  dat <- tibble(
    c = rnorm(n),
    x = rnorm(n, eff_cx * c, 1),
    y = rnorm(n, eff_cy * c + eff_xy * x, 1)
  )
  if (controlForConfound) {
    model <- lm(y ~ x + c, data = dat)
  } else if (!controlForConfound) {
    model <- lm(y ~ x, data = dat)
  }

  tidy(model) %>% filter(term == "x") %>% pull(estimate)
}

coefs <- replicate(n, simulate_study())

```

As we can see, the median effect is indistinguishable from 0; in other words, the confounding has completely masked the true .2 effect.

```{r}
#| label: plot-results

hist(coefs)
abline(v = median(coefs), col = "red", lwd = 4)
```

If we run the same thing again but include the confound in the model, we recover the true .2 effect.

```{r}
#| label: control-for-confound

unbiasedCoefs <- replicate(n, simulate_study(controlForConfound = TRUE))
hist(unbiasedCoefs)
abline(v = median(unbiasedCoefs), col = "blue", lwd = 4)
```